---
title: "exercise2"
author: "Ian Nel"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 2: Explore two variables with dplyr and ggplot2

## Loading Packages
```{r}
library(ggplot2)
library(dplyr)
library(palmerpenguins)
```

## 2.1 
Pick two quantitative variables to explore.
* Make a summary table of descriptive statistics for these variables using summarize(). Include whatever staistics you feel appropriate (mean, median sd, range, etc.).
* Make a scatterplot of these variables using ggplot().

### 2.1a)
```{r}
penguins %>% 
  summarize(mu_flipper_length = mean(flipper_length_mm, na.rm = TRUE),
            sigma_flipper_length = sd(flipper_length_mm, na.rm = TRUE),
            mu_body_mass = mean(body_mass_g, na.rm = TRUE),
            sigma_body_mass = sd(body_mass_g, na.rm = TRUE))
            
```

### 2.1b)
```{r}
penguins %>% 
  ggplot(aes(x=flipper_length_mm, y=body_mass_g)) +
  geom_point(alpha = .4) +
  theme_bw()
```

## 2.2

Pick one categorical variable and one quantitative variable to explore.

* Make a summary table giving the sample size (hint: n()) and descriptive statistics for the quantitative variable by group.

* Make one or more useful plots to visualize these variables.

### 2.2a)
```{r}
penguins %>% 
  group_by(species) %>% 
  summarize(mu_bill_length = mean(bill_length_mm, na.rm = TRUE),
            sigma_bill_length = sd(bill_length_mm, na.rm = TRUE),
            n = n())

```

### 2.2b) 
```{r}
penguins %>%  
  ggplot(aes(bill_length_mm, species)) +
  ggridges::geom_density_ridges()
```