---
title: "hw_2"
author: "Ian Nel"
date: "2/8/2021"
output: 
  html_document: 
    theme: paper
  github_document: default
---

<!-- BMW: Make the md version of your document this way. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2) # BMW: Put the library() calls in the `setup` chunk
library(dplyr)
library(gapminder)
```

# Exercise 1: Basic dplyr

Use dplyr functions to achieve the following exercises 

<!-- BMW: You need blank lines between paragrahs in markdown. A single new line is treated as a space. -->

### Gapminder dataset 

```{r}
gapminder #BMW: Don't print out how datasets like this in your final reports.
```

## 1.1

Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.

```{r}
gapminder %>% 
  filter(year > 1969 & year < 1980 & country %in% c("Afghanistan", "Albania", "Algeria")) # BMW: Great!
```

## 1.2 

Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1

```{r}
gapminder %>% 
  filter(year > 1969 & year < 1980 & country %in% c("Afghanistan", "Albania", "Algeria")) %>% 
  select(country, gdpPercap) #the c() is superfluous BMW: specifically in select()
```

## 1.3

Make a new variable in gapminder for the change in life expectancy from the previous measurement. 
Filter this table to show all of the entries that have experienced a drop in life expectancy.

Hint: you might find the lag() or diff() functions useful.

```{r}
gapminder %>% 
  group_by(country) %>%  
  mutate(delta_lifeExp = lifeExp - lag(lifeExp)) %>% 
  filter(delta_lifeExp < 0) # BMW: Great!
```

## 1.4 

Filter gapminder so that it shows the max GDP per capita experienced by each country.

Hint: you might find the max() function useful here.

```{r}
gapminder %>% 
  group_by(country) %>% 
  filter(gdpPercap==max(gdpPercap))
```

## 1.5

Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. In your plot, put GDP per capita on a log scale.

```{r}
gapminder %>% 
  filter(country=="Canada") %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point(alpha = 0.5) +
  scale_x_log10() +
  ggtitle("Canada’s life expectancy vs. GDP per capita") +
  theme_bw()
```


