---
title: "2_11_21_Notes"
author: "Ian Nel"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(palmerpenguins)
library(gapminder)

```


## ggplot
```{r}
penguins %>%  
  mutate(Sex = recode(sex, 
                        female = "Female", 
                        male = "Male")) %>% 
  ggplot() +
  aes(x = species, y = flipper_length_mm,
      color = Sex) +
  geom_jitter(height = 0, width = .2) + 
  xlab("Species") +
  scale_y_continuous("Flipper Length (mm)") +
  scale_color_brewer(palette = "Accent") +
  theme_classic()
  


```
x-lab vs. scale_x_discrete 
x_lab - just change the title 
labs - does the same thing as x_lab 
scale functions - general utility knife for doing all sorts of things; other formatting 

brewer is good for people with color deficiencies

aes - mapping
geom - shape
scale - controlling appearance

## BFI rescoring

```{r}
# scales - A C E N O 
# 5 each 

bfi_scored <- #make new data frame with my scored scores 
  psych::bfi %>% 
  rowwise() %>% 
  mutat() #compute mean score for each row

```

#Dplyr review
https://talks.andrewheiss.com/2020-gpl/01_welcome-tidyverse/04_transform-data.html

%in% Test each row against all values in set instead of testing each 
relationalfunctions = element-wise comparison (each element against one other elemtn in comparison vector)
* compares input vector to test values 
```=``` assignment; ```==``` equal to 
```" "``` character values (e.g., "Canada" vs. Canada)
"," & "&" do the same thing "|" = or 
! excludes 
```{r}
gapminder %>% 
  mutate(pop_mil = round(pop / 1000000)) #rounds to nearest whole number (in millions)


```

ifelse(Test, value if true, value if false)
* Test = logical test 
* value_if_true = what happens if it's true 
* value_if_false = what happens if it's false 
```{r}
mutate(gapminder, 
       after_1960 = ifelse(year > 1960))

```

## summarize
```{r}
gapminder %>% 
  summarize(mean_life = mean(lifeExp)) #name of column, computed as mean of lifeExp

gapminder %>% 
  filter(continent == "Africa", year == 2007) %>% #only africa in year 2007
  summarize(n_countries = n_distinct(country),
            med_le = median(lifeExp))

```


```{r}


```


```{r}

```


