---
title: "2.9.21"
author: "Ian Nel"
date: "2/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages
```{r}

library(dplyr)

```

# In-Class Exercise
 Use the `psych::bfi` dataset.
2. Compute mean scores for each of the Big Five traits for each person.
3. Keep just Big Five scores, gender, age, and education.
4. Summarize the mean, standard deviation, minimum, and maximum values for the Big Five scores separately by gender groups.
5. Summarize the mean, standard deviation, minimum, and maximum values for the Big Five scores separately by educational level groups.

## 2 & 3 
```{r}
bfi_scored <- psych::bfi %>% 
  rowwise() %>% 
  mutate(A_tot = mean(c_across(A1:A5), na.rm = TRUE),
         C_tot = mean(c_across(C1:C5), na.rm = TRUE),
         E_tot = mean(c_across(E1:E5), na.rm = TRUE),
         N_tot = mean(c_across(N1:N5), na.rm = TRUE),
         O_tot = mean(c_across(O1:O5), na.rm = TRUE)) %>% 
  select(gender:O_tot)

```

## 4 
Summarize the mean, standard deviation, minimum, and maximum values for the Big Five scores separately by gender groups.
```{r}
bfi_scored %>% 
  group_by(gender) %>% 
  summarize(
    across(A_tot:O_tot, #don't need c because I'm using the ":"
           list(mu = mean, SD = sd, minimum = min, maximum = max)
    )
  )

```


## 5 
Summarize the mean, standard deviation, minimum, and maximum values for the Big Five scores separately by educational level groups
```{r}
bfi_scored %>% 
  group_by(education) %>% 
  summarize(
    across(A_tot:O_tot, #don't need c because I'm using the ":"
           list(mu = mean, SD = sd, minimum = min, maximum = max)
    )
  )

```
```{r}


```


## Summarize() practice

```{r}

gapminder::gapminder %>% 
  summarize(mean_pop = mean(pop),
            mean_gdp = mean(gdpPercap)) 
  
gapminder::gapminder %>% 
  summarize(across(c(pop, gdpPercap), list(mn = mean, SD = sd))) #streamlined version of code above
            
gapminder::gapminder %>% 
  group_by(continent) %>% 
  summarize(across(c(pop, gdpPercap)), #across() means do this thing across multiple different columns  
                   list(mn = mean, SD = sd))
            
gapminder::gapminder %>% 
  group_by(continent) %>% 
  summarize(across(c(pop, gdpPercap), 
                   list(mn = ~ mean(.x, na.rm = TRUE), SD = sd)) #tilda brings in whole function; .x "that column"
```



rowwise function think of as being group_by (row ) treats each row as a separate group; use to calculate stuff on multiple items
```{r}
psych::bfi %>% 
  rowwise() %>% 
  mutate(A_tot = mean(c(A1, A2, A3, A4, A5), na.rm = TRUE),
        (C_tot = mean(c_across(C1:C5), na.rm = TRUE)) %>% 
  select(A_tot, C_tot)

```

```{r}


```
