---
title: "Notes_3/2/21_File_Input/Output"
author: "Ian Nel"
date: "3/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

csv = comma separated values; great format for exchanging data
Save excel files as csv before further processing because it's simpler

install.packages("here")
here package best way to refer to files - provides sequence of folders to follow; sned them entire R projects
here::here()

## loading packages
```{r}
library(tidyverse)

```

## Here:here example - finding README.md in hw1 file
```{r}
here::here("participation","README.md") #first argument is where is the path/file starts, 

```
## herehere - reading in green data file in h3 folder, then extracting column specifications
```{r}
read_csv(
  here::here("hw_3","green_data.csv")) %>% 
  spec()

```
copy and paste column specifications from ```read_csv``` to col_types; read_csv guesses and does a pretty good job; saves time so you don't have to manually type column specifications

na = c("", "NA") defaults to 2 different things; NA = blank; NA = "NA"; can add more NA specifications (e.g., c("","NA",-999))

recursive = make all of the necessary folders along the way 

## Here here function 
```here::here``` list out sequence of folders (in quotations) you want to jump into, separated by commas
