---
title: "hw3"
author: "Ian Nel"
date: "2/22/2021"
output: 
  github_document: default 
  html_document:
    theme: cerulean
---

<!-- BMW: In your YAML header, indent subsections by 2 spaces, not 4.
          Sub-options, like the theme for html_document, need to be indented under the parent option. When you have a sub-option like theme, don't include "default" for the parent option. 
          See what I edited above.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reshape2) # BMW: Don't use reshape2 and tidyverse packages (eg, dplyr or tidyr) together in the same script. reshape2 isn't supported anymore, and I suggest moving to using tidyr::pivot_longer() and tidyr::pivot_wider() like we practiced in class instead.
library(kableExtra)
library(scales)
library(ggplot2)
```

# Homework 03

You will analyze data looking at the relationship between **green reputation** and three personality traits--**compassion**, **intellectual curiosity**, and **openness to experiences**.
The dataset includes data from **students** and **non-students**.


```{r data}
dictionary <- readr::read_csv(here::here("hw_3", "green_dictionary.csv"))
green_data <- readr::read_csv(here::here("hw_3", "green_data.csv"))
```

For your assignment, do the following.

  1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). 
     Is anything unusual?
  2. Compute total scores for the four scales. 
     Recode variables as needed.
  3. Rescale the variables so that they go from 0-100 instead of the original range.
     Name the recaled variables `*_pomp`.
  4. Make plots that illustrate the distributions of the 4 POMP-scored variables.
  5. Make scatterplots showing the relationships between **green reputation** and each personality trait.
     Include trend lines for **students** and **non-students**.
     What do these plots show?
  6. Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).
  7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.
  
In your assignment, prepare an RMarkdown file that includes both the coding and (for 4-7) narrative descriptions of your results.

## 1) Inspect the item responses (e.g., with graphs or by summarizing distinct values). Is anything unusual?

```{r item inspection}
print(green_data) # BMW: Dont' print whole datasets in your final report
summary(green_data) #some values are coded as -99, so I'm going to code them as

dat <- green_data %>%
  mutate(student = factor(student, levels = c(1, 2), labels = c("Non-student", "Student"))) %>% # BMW: You can conversion to factor and setting labels in one line like this
  na_if("-99")

summary(dat) #now minimum values are 1 (not -99)
```

## 2) Compute total scores for the four scales. Recode variables as needed.
```{r}
#see which variables need to be recoded
view(dictionary) # BMW: Don't leave `View()` calls in your final document. 

#Reverse code items
reversed <- dictionary %>% 
  filter(Keying == -1) %>% 
  pull(Item)

datr <- dat %>% 
  mutate(
    across(all_of(reversed),
           ~ 6 - .x,
           .names = "{.col}r")
  ) # BMW: Nicely done.

#Using rowwise function to create total scores for each person 

datr2 <- datr %>% 
  rowwise() %>% 
  mutate(
   green_tot = mean(c(green1,green2r,green3,green4,green5), na.rm = FALSE),
         comp_tot = mean(c(comp1,comp2,comp3,comp4,comp5,comp6r,comp7r,comp8r,comp9r,comp10r), na.rm = FALSE),
         intel_tot = mean(c(intel1,intel2,intel3,intel4,intel5,intel6,intel7r,intel8r,intel9r,intel10r), na.rm =FALSE),
         open_tot = mean(c(open1,open2,open3,open4,open5,open6,open7r,open8r,open9r,open10r), na.rm = FALSE)
   ) %>% 
  ungroup() # BMW: Always ungroup() after you use rowwise() or group_by()

summary(datr2)
```

##  3) Rescale the variables so that they go from 0-100 instead of the original range. Name the recaled variables `*_pomp`.

<!-- BMW: You need a blank line between text and code chunks to ensure correct printing when knit. -->

```{r}
#Using rowwise function to rescale scores for each person from a 1-5 to a 0-100 scale. 
datr3 <- datr2 %>% 
  rowwise() %>% 
  mutate(green_pomp = rescale(green_tot, to = c(0, 100), from = c(1,5)),
         comp_pomp = rescale(comp_tot, to = c(0, 100), from = c(1,5)),
         intel_pomp = rescale(intel_tot, to = c(0, 100), from = c(1,5)),
         open_pomp = rescale(open_tot, to = c(0, 100), from = c(1,5))
  ) %>% 
  ungroup()

# BMW: In general, use more descriptive object names. eg, dat_cleaned or dat_analysis instead of datr3

```

## 4) Make plots that illustrate the distributions of the 4 POMP-scored variables.

```{r}
green_hist <- ggplot(datr3) +
  aes(x = green_pomp, y = ..density..) +
  geom_histogram(color="springgreen",fill="dark green", binwidth = 5)+
  geom_density(alpha = .1, fill = "black")

comp_hist <- ggplot(datr3) +
  aes(x = comp_pomp, y = ..density..) +
  geom_histogram(color = "midnightblue", fill="pink", binwidth = 5)+
  geom_density(alpha = .1, fill = "black")

intel_hist <- ggplot(datr3) +
  aes(x = intel_pomp, y = ..density..) +
  geom_histogram(color = "darkorchid2", fill="gold", binwidth = 5) +
  geom_density(alpha = .1, fill = "black")

open_hist <- ggplot(datr3) +
  aes(x = open_pomp, y = ..density..) +
  geom_histogram(color="black", fill="turquoise2", binwidth = 5) +
  geom_density(alpha = .1, fill = "black")

gridExtra::grid.arrange(green_hist,comp_hist,intel_hist,open_hist)

# BMW: Nicely done. 
```
### Narrative Description 

These histograms demonstrate that these variables have a relatively normal distribution, except for "comp_pomp" which is skewed to the left. "Intel_pomp" also appears to be somewhat bimodal. 

## 5) Make scatterplots showing the relationships between **green reputation** and each personality trait. Include trend lines for **students** and **non-students**. What do these plots show?

```{r}
datr3 %>% 
  filter(!is.na(student)) %>% 
  ggplot() +
  aes(y = comp_pomp,
      x = green_pomp,
      group = student,
      color = student,
      fill = student) +
  geom_jitter() +
  geom_smooth(method = "lm")
  
datr3 %>% 
  filter(!is.na(student)) %>% 
  ggplot() +
  aes(y = intel_pomp,
      x = green_pomp,
      group = student,
      color = student,
      fill = student) +
  geom_jitter() +
  geom_smooth(method = "lm")
  
datr3 %>% 
  filter(!is.na(student)) %>% 
  ggplot() +
  aes(y = open_pomp,
      x = green_pomp,
      group = student,
      color = student,
      fill = student) +
  geom_jitter() +
  geom_smooth(method = "lm")

```
### Narrative Description

<!-- BMW: You need a blank line between paragraphs. -->

There is a small positive correlation between green_pomp and all three personality traits. The strength and magnitude of these correlations are similar for both students and nonstudents, but the relationship between green_pomp and comp_pop is slightly greater for students than non-students.

## 6) Compare **green reputation** for students and non-students using a **rainfall plot** (bar + density + data points).

```{r}
datr3 %>%
  filter(!is.na(student)) %>%
  ggplot() +
  aes(y = student,
      x = green_pomp,
      fill = student,
      color = student) + 
  geom_jitter(width = 0, height = .1) + # BMW: Rainfall plots include the raw data points too
  geom_boxplot(
    color = "black",
    alpha = .5,
    width = .5,
    size = .5
  ) +
  ggdist::stat_slab(
    height = .5,
    color = "black",
    size = .5,
    alpha = .5,
    position = position_nudge(y = .1)
  )
```

### Narrative Description 

This rainfall plot shows that the median green_pomp for non-students is slightly larger for non-students than for students. The green_pomp distribution for non-students appears to be more bimodal, whereas the distribution for students appears to be normal.

## 7) Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.

```{r}
sumtable <- datr3 %>% 
  na.omit() %>% 
  group_by(student) %>% 
  summarize(across(c(green_pomp,open_pomp,intel_pomp,comp_pomp),
                   list(Mean = ~ mean(.x, na.rm = TRUE),
                        SD = ~ sd(.x, na.rm = TRUE),
                        Median = ~ median(.x, na.rm = TRUE),
                        Min = ~ min(.x, na.rm = TRUE),
                        Max = ~ max(.x, na.rm = TRUE)))
  )
kable(sumtable)

```

Non-students had a greater mean and median green reputation and intellectual curiosity compared with students, whereas students had a higher mean and median openness to experience and compassion. 
