---
title: "Beer Advocate: ABV and quality ratings"
author: "Your Name Here"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Packages 
```{r}
library(tidyverse)
library(readr)
```

For this activity, download the data file at this location to the `data` subfolder in your `participation` folder. 
This dataset contains information about different catagories of beer as rated on <beeradvocate.com>.
The data are originally from [here](http://scottjanish.com/relationship-of-abv-to-beer-scores/).


The dataset contains 3 variables:

  1. Beer category
  2. Average alcohol by volume (ABV) in percentage (%)
  3. Average quality rating (rAVG) on a 1–5 scale
  
You should write a script to do the following:

  1. Download the dataset with an appropriate filename.
  2. Import the dataset into R.
     Be sure to check for any importing problems and correct as needed.
  3. Make a figure illustrating the relationship between ABV and beer quality.
  4. Compute a summary table with the mean and SD of each variable, along with their correlation.
  5. Describe the results.

## 1 Download with filename
```{r}
if ( !dir.exists( here::here("participation", "data") ) ) {
  dir.create( here::here("participation", "data"), recursive = TRUE )
} #if this doesn't exist, then create this "data" directory/folder

url <- "https://raw.githubusercontent.com/USF-Psych-DataSci/Classroom/master/data/beer-advocate.csv"
download.file(
  url, 
  here::here("participation", "data", "beerdata.csv"), 
  mode = "wb"
)
```

##2 Import to R
```{r}
beerdata <- read_csv(
  here::here("participation", "data", "beerdata.csv"),
  col_types = cols(
    `Beer Category` = col_character(),
    `Average ABV` = col_double(),
    rAVG = col_double()
  ),
  na = c("", "NA", "Non-alcoholic")
)
#always use read_csv not read.csv; read_csv won't rename your columns & checks type of variable in each column
#ctrl-i re-indents the lines; check out code tab
#If i want to tell what to treat things as missing, use ```na``` argument to specify
```

## 3 Make a figure
```{r}
beerdata %>% 
  ggplot() +
  aes(`Average ABV`, rAVG) +
  geom_point(alpha = .6) +
  stat_smooth(method=lm) +
  theme_classic()
```

## 4 Compute a summary table with the mean and SD of each variable, along with their correlation.
```{r}
beerdata %>% 
  summarise(across(c(`Average ABV`, rAVG),
            list(
              mean = ~mean(.x, na.rm= TRUE),
              SD = ~sd(.x, na.rm = TRUE),
              r = cor.test(`Average ABV`, rAVG) # this is wrong
            )))

```

## 5 describe results 